---
title: "COVID-19 Tracker Indonesia"
author: "[www.datawizart.com](https://datawizart.com/)" 
date: "based on data: 7/04/2020"
output:
  rmarkdown::html_document:
    theme: simplex
    code_folding: "hide"
    template: "template_covid19.html"
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}

h2.subtitle{
  font-size: 20px;
  color: DarkRed;
  text-align: center;

}
h4.author { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  text-align: center;
}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.retina=1.2)
```

```{r library, include=FALSE}
library(tidyverse)
library(geofacet) 
library(ggrepel)
library(plotly)
library(gghighlight)
library(RcppRoll)
library(cowplot)
library(ggtext)
source("global_function.R")
```

```{r load data indonesia, include=FALSE}
df_indo <- read_rds("df_indo.rds")
indo_grid <- read.csv("~/indo_grid.csv",sep =  "\t",stringsAsFactors = FALSE)

```

```{r mempersiapkan dataset, include=FALSE}
date_variable <- "2020-07-04"

heatmap_indo <- df_indo %>%
  arrange(data_date) %>% 
  group_by(province) %>%
  mutate(casesroll_avg=roll_mean(daily_case, 7, align="right", fill=0)) %>% 
  mutate(maxcaserate=max(casesroll_avg), 
         maxcaseday=data_date[which(casesroll_avg==maxcaserate)][1]) %>% 
  ungroup()

heatmap_indo$province <- fct_reorder(heatmap_indo$province, -1*heatmap_indo$total_case,min)


filter_daily <- heatmap_indo %>%
  group_by(province) %>% 
  slice(min(which(casesroll_avg>=5))) %>% 
  ungroup() %>% 
  rename(filter_week = data_date)

heatmap_daily_indo <- heatmap_indo %>% ungroup() %>% 
  left_join(filter_daily[,c(1,2)], by = c("province")) %>% 
  filter(data_date >= filter_week) %>% 
  arrange(data_date) %>% 
  group_by(province) %>% 
  mutate(seq_ = row_number()) %>% ungroup()

#heatmap_daily_indo$province <- 
#  fct_reorder(.f = (heatmap_daily_indo$province),.x = #-1*heatmap_daily_indo$total_case,.fun = min)
```

### What is the condition of your province? Has it passed the peak?

To answer this question we made three graphs, each graph can be used to analyze Covid-19 cases in 34 provinces in Indonesia.
These three graphs display a 7-day rolling average of the number of positive Covid-19 cases in each province. The first and second graphs are replications of the dashboard created by **Financial Times** [link](https://ig.ft.com/coronavirus-chart/?areas=usa&areas=idn&cumulative=0&logScale=1&perMillion=0&values=cases). 
While the third graph is the heatmap replication of the account **twitter @VictimOfMaths** [source](https://github.com/VictimOfMaths/COVID-19).
These three graphs complement each other, there are advantages and disadvantages in each graph. By displaying all three hopefully, conclusions can be drawn.

#### Time-series Rolling Average New Positive Case

NOTE:
In this first graph, we map the rolling-averages of one province in one time-series line so it is easier to do comparisons between provinces. In this graph, rolling averages begin when the number of cases has reached 4 or more cases.

```{r rolling avg kasus positif,fig.align='center',fig.height=9,fig.width=10}
heatmap_daily_indo %>% group_by(province) %>% 
  mutate(dott_y = if_else(seq_==max(seq_),casesroll_avg,NaN)) %>% 
  ungroup() %>% 
  ggplot(aes(x=seq_ ,group=province ))+
  geom_line(aes( y = casesroll_avg))+
  geom_point(aes(y = dott_y),size=1)+
  gghighlight(use_direct_label = FALSE,unhighlighted_params = list(size=0.4))+
  scale_y_continuous(trans='log2',
                     breaks=c(1,10,100,200,500,1000))+
  facet_wrap(~province)+
  labs(subtitle = "7 days rolling average positif cases, since # of positive cases above 4",
       y = "Log Rolling Avg Positif Case",
       x = "Number of days from since # of positive cases above 4",
       caption = paste("Code by Dio Ariadi | www.datawizart.com. Graph inspired by twitter @jburnmurdoch\nData Source: kawalCOVID19 | covid19.go.id",date_variable))+
  theme(panel.background = element_rect(fill = "#f5f5f5"),
        text = element_text(family = "Proxima Nova"),
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "#f5f5f5"),
        plot.background = element_rect(fill = "#f5f5f5"),
        strip.text.x = element_text(hjust = 0),
        plot.caption = element_text(hjust = 0))

```

#### Time-series Positive Daily Cases and 7 Days Rolling Average

In the second graph, we can see in more detail the increase in cases in each province. In addition to using a line chart that illustrates the rolling average, this chart uses a bar chart that illustrates the number of positive cases per day in the province. We can see clearly on what day the peak occurred in a province.

```{r Time-series Kasus Positif Harian dan Rolling Average 7 Hari,fig.align='center',fig.height=8,fig.width=10}
heatmap_indo %>% filter(total_case>0 & province!="no_define") %>% 
  arrange(data_date) %>% 
  group_by(province) %>% 
  mutate(seq_=row_number()) %>% 
  ungroup() %>% 
  ggplot(aes(x=seq_))+
  geom_col(aes(y = daily_case),alpha=0.5)+
  geom_line(aes(y = casesroll_avg))+
  labs(title = "# of daily positive cases and 7-day rolling averages",
       subtile = "Bars are daily positive cases, lines are rolling avg",
       x = "Days from the first positive case",
       y = "# Positive Case",
       caption = paste("Code by Dio Ariadi | www.datawizart.com\nData Source: kawalCOVID19 | covid19.go.id",date_variable))+
  facet_wrap(~province,scales = "free",ncol = 5)+
  theme(panel.background = element_rect(fill = "#f5f5f5"),
        text = element_text(family = "Proxima Nova"),
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_rect(fill = "#f5f5f5"),
        plot.background = element_rect(fill = "#f5f5f5"),
        strip.text.x = element_text(hjust = 0),
        plot.caption = element_text(hjust = 0))
```

#### Heatmap Each Province

This heatmap illustrates the 7 day rolling average of normalized positive cases based on the largest positive case value in each province. The greener the map of a province means the further the rolling average moves from the highest number of positive cases. This can be interpreted as saying that the province has passed the peak of new cases.

```{r heatmap kasus positif,fig.align='center', fig.height=10, fig.width=12, message=FALSE, warning=FALSE, dpi=60}
heatmap_indo_2 <- heatmap_indo %>% filter(province!="no_define")
heatmap_indo_2$province <- fct_reorder(heatmap_indo_2$province, heatmap_indo_2$total_case,max)

#https://github.com/VictimOfMaths/COVID-19
heatmap_indo_2$maxcaseprop <- heatmap_indo_2$casesroll_avg/heatmap_indo_2$maxcaserate
casetiles <- ggplot(heatmap_indo_2, aes(x=data_date, y=fct_reorder(province, total_case,max), fill=maxcaseprop))+
  geom_tile(colour="White", show.legend=FALSE)+
  theme_classic()+
  scale_fill_distiller(palette="Spectral")+
  scale_y_discrete(name="", expand=c(0,0))+
  
  labs(title = "Heatmap timeline daily additional positive cases",
       subtitle = "The color of each day is compared with the highest day",
       caption=paste("Code by @VictimOfMaths | Replicated & modified by Dio Ariadi | www.datawizart.com\nData Source: kawalCOVID19 | covid19.go.id",date_variable))+
  theme(
    axis.line.y=element_blank(), 
    plot.subtitle=element_text(size=rel(0.78)), 
    plot.title.position="plot",
    axis.text.y=element_text(colour="Black"),
    text = element_text(family = "Proxima Nova"),
    plot.background = element_rect(fill = "#f5f5f5"),
    panel.background = element_rect(fill = "#f5f5f5"),
    plot.caption = element_text(hjust = 0),
    plot.caption.position = "plot"
    )

casebar <- ggplot(subset(heatmap_indo_2, data_date==max(data_date)), aes(x=total_case, y=fct_reorder(province, total_case,max), fill=total_case))+
  geom_col(show.legend=FALSE)+
  theme_classic()+
  scale_fill_distiller(palette="Spectral")+
  scale_x_continuous(name="Cummulative Positive Cases", breaks=c(0,1000,5000,8000))+
  theme(axis.title.y=element_blank(), 
        axis.line.y=element_blank(), 
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), 
        axis.text.x=element_text(colour="Black"),
        text = element_text(family = "Proxima Nova"),
        plot.background = element_rect(fill = "#f5f5f5"),
        panel.background = element_rect(fill = "#f5f5f5"))

plot_grid(casetiles, casebar, align="h", rel_widths=c(1,0.2))
```

### Growth of Confirmed Case

In graph Four, we display a cumulative log graph of the number of cases from the day the number of cases is more than 4. The dashed line guides to see how the acceleration of the addition of positive cases in each city. If the solid black line is approximately has same trajectory with some of the dashed lines it means that the case will double as the same as the value of dashed line.

```{r pergerakan kasus positif,fig.align='center',fig.height=8,fig.width=10}
label_case=
  bind_rows(data.frame(value = seq_fun(start_value = 5, end_value = max(df_indo$total_case), 3), double_each = 3) %>% mutate(seq_ = row_number()-1),
            data.frame(value = seq_fun(start_value = 5, end_value = max(df_indo$total_case), 2), double_each = 2) %>% mutate(seq_ = row_number()-1),
            data.frame(value = seq_fun(start_value = 5, end_value = max(df_indo$total_case), 7), double_each = 7) %>% mutate(seq_ = row_number()-1),
            data.frame(value = seq_fun(start_value = 5, end_value = max(df_indo$total_case), 4), double_each = 4) %>% mutate(seq_ = row_number()-1),
            data.frame(value = seq_fun(start_value = 5, end_value = max(df_indo$total_case), 14), double_each = 14) %>% mutate(seq_ = row_number()-1))

label_case <- label_case %>% pivot_wider(names_from = double_each,values_from = value)


color_label <-  c("Case Doubled each 3 days"="grey","...4 days"="red","...7 days"="darkred","...14 days"="blue")
case_cumulative <- heatmap_indo %>% filter( total_case>4 & province!="no_define") %>% 
  group_by(province) %>% 
  arrange(data_date) %>% 
  mutate(seq_ = row_number())  %>% 
  ungroup() %>% 
  group_by(province,total_case) %>% 
  mutate(seq_1 = row_number()) %>% ungroup()
case_cumulative <- case_cumulative %>% left_join(label_case,by = c("seq_"))
case_cumulative <- case_cumulative %>% 
  group_by(province) %>% 
  mutate(`3` = ifelse(`3`>max(total_case),NA,`3`),
         `2` = ifelse(`2`>max(total_case),NA,`2`),
         `7` = ifelse(`7`>max(total_case),NA,`7`),
         `4` = ifelse(`4`>max(total_case),NA,`4`),
         `14` = ifelse(`14`>max(total_case),NA,`14`))
case_cumulative %>% 
  ggplot()+
  geom_line(aes(x=seq_,y=total_case) )+
  geom_line(aes(x = seq_,y = `3`) ,linetype="dashed",color="grey")+
  geom_line(aes(x = seq_,y = `14`),linetype="dashed",color="blue")+
  geom_line(aes(x = seq_,y = `4`) ,linetype="dashed",color="red")+
  geom_line(aes(x = seq_,y = `7`) ,linetype="dashed",color="darkred")+
  scale_y_continuous(trans='log2',
                     breaks=c(10,100,500,1000,2000,5000,10000,25000,80000,120000))+
  facet_wrap(~province,scales = "free")+
  scale_color_manual(values = color_label)+
  labs(y = "Log total positive csae", x = "Days since # of positive case is 5",
       caption = paste("Code by Dio Ariadi & Nur Izzahudin | www.datawizart.com\nGraph inspired by @jburnmurdoch\nData Source: kawalCOVID19 | covid19.go.id",date_variable),
       title = "Cummulative positive cases, since 5 positive cases",
       subtitle = "<br><span style='color:grey'>Dashed Line (Case doubled, each 3 days)</span> <span style='color:red'>Dashed Line (. . .4 days)</span> <span style='color:darkred'>Dashed Line (. . . 7 days)</span> <span style='color:blue'>Dashed Line (. . .14 days)</span>",color="legend")+
  theme(panel.background = element_rect(fill = "#f5f5f5"),
        text = element_text(family = "Proxima Nova"),
        panel.grid.major = element_line(colour = "#f0f0f0"),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(size = 6),
        strip.background = element_rect(fill = "#f5f5f5"),
        plot.background = element_rect(fill = "#f5f5f5"),
        strip.text.x = element_text(hjust = 0),
        plot.subtitle = element_markdown(size = 8),
        plot.caption = element_text(hjust = 0))
  
```


### Death Rate Each Province

The death rate is defined as the number of death cases divided by the number of positive cases. We can see that the provinces with the highest mortality rates are mostly found on the islands of Java and Sumatra.


```{r ratio kematian per provinsi, fig.align='center', fig.height=7, fig.retina=1.2, fig.width=12, message=FALSE, warning=FALSE}
death_ratio_df <- df_indo %>% 
  group_by(code,province) %>% 
  arrange(data_date) %>% 
  mutate(rank_ = row_number(),
         n_ = n()) %>% 
  filter(rank_ == n_) %>% 
  ungroup() %>% 
  mutate(death_ratio = total_death/total_case,
         death_ratio = ifelse(is.na(death_ratio),0,death_ratio)) %>% 
  left_join(indo_grid[,c(1,2,5,6)], by = c("province"="name_indo")) 

death_ratio_df %>% 
  ggplot(aes(x = (col), y = -1*row,
             fill = death_ratio))+
  geom_tile(color = "#f5f5f5",size=1)+
  geom_text(aes(label=round(death_ratio,3)*100),vjust=1.2,family = "Proxima Nova")+
  geom_text(aes(label=name_short),vjust =-1,family = "Proxima Nova",size = 3)+
  labs(x = "",y= "", caption = paste("Created by Dio Ariadi\nwww.datawizart.com\nData Source: kawalCOVID19 | covid19.go.id",date_variable))+
  coord_equal()+
  scale_fill_distiller(palette="Spectral")+
   theme(panel.grid = element_blank(),
        text = element_text(family = "Proxima Nova"),
        strip.background = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text.x = element_text(size = 7),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.background=element_rect(fill = "#f5f5f5"),
        panel.background = element_rect(fill = "#f5f5f5"),
        legend.position = "none",
        plot.caption = element_text(hjust = 0))


```

```{r bar plot death ration, fig.align="center"}
overall_indo <- sum(death_ratio_df$total_death)/sum(death_ratio_df$total_case)
overall_world <- read_rds("df_worldmeter_analysis.rds") %>% 
  group_by(country) %>% slice(which.max(data_date)) %>% ungroup() %>% summarise(sum(deaths,na.rm = TRUE)/sum(cases,na.rm = TRUE)) %>% pull()



death_ratio_df %>% filter(province!="no_define") %>% 
  ggplot(aes(x = fct_reorder(province,death_ratio),
             y = death_ratio,
             fill = death_ratio))+
  geom_col(color = "#f5f5f5",size=1)+
  geom_hline(aes(yintercept= overall_indo),linetype ="dashed",color = "darkred")+
  geom_hline(aes(yintercept= overall_world),linetype ="dashed",color = "darkblue")+
  #geom_text(aes(x = 4, y = 0.082, label = paste("Rasio Kematian\n Indonesia:",round(overall_indo,3)*100)), family= "Proxima Nova",size =3)+
  annotate("text", x = 4, y = ifelse(overall_indo>overall_world,overall_indo+0.003,overall_indo-0.003), label = paste("Death Rate\nIndonesia:",round(overall_indo,3)*100,"%"),family = "Proxima Nova",
           size=3,
           colour ="darkred",
           hjust = ifelse(overall_indo>overall_world,0,1))+
    annotate("text", x = 8, y = ifelse(overall_world > overall_indo,overall_world+0.003,overall_world+0.003), label = paste("Death Rate\nWorldwide:",round(overall_world,3)*100,"%"),family = "Proxima Nova",
           size=3,
           colour ="darkblue",
           hjust = ifelse(overall_indo>overall_world,1,0))+
  labs(x = "",
       y = "Death Ratio",
       caption = paste("Created by Dio Ariadi | www.datawizart.com\nData Source: kawalCOVID19 | covid19.go.id",date_variable))+
  coord_flip()+
  scale_fill_distiller(palette="Spectral")+
  scale_y_continuous(labels = scales::percent)+
   theme(panel.grid = element_blank(),
        text = element_text(family = "Proxima Nova"),
        strip.background = element_blank(),
        axis.ticks.x = element_line(colour = "grey"),
        strip.text.x = element_text(size = 7),
        axis.ticks.y = element_line(colour = "grey"),
        plot.background=element_rect(fill = "#f5f5f5"),
        panel.background = element_rect(fill = "#f5f5f5"),
        legend.position = "none",
        plot.caption = element_text(hjust = 0),
        plot.caption.position =  "plot")
  #geom_text(aes(label=round(death_ratio,3)*100),vjust=1.2,family = "Proxima Nova")+
  #geom_text(aes(label=name_short),vjust =-1,family = "Proxima Nova",size = 3)
```

### Can it be used as a basis for decisions?

My answer will depend on your opinion on how much do you trust the government data.
As a reference [Jakarta deaths rose 52% compared to the previous year's average](https://twitter.com/jburnmurdoch/status/1260233704141619205?s=20). This is as a consideration whether the data can be used to infer something and the data above does not take into account the number of tests conducted on the community.
Again all this data is not just a number, but someone life.

### Feedback and Suggestion
I also provide the code on github to help other analysts and appreciate if the source can be included as a citation.

- email `dioariadi11@gmail.com`


### Appendix

#### Contributor

LinkedIn:

- [Dio Ariadi](https://www.linkedin.com/in/dioariadi/)
- [Nur Izzahudin](https://www.linkedin.com/in/nur-izzahudin/)
- [Hamidah Alatas](https://www.linkedin.com/in/hamidah-alatas-4a5675126/)


#### PSBB Date

- [Jakarta PSBB 10 April](https://megapolitan.kompas.com/read/2020/04/22/18150051/psbb-jakarta-resmi-diperpanjang-28-hari-hingga-22-mei-2020)
- [Bandung PSBB 22 April](https://www.cnnindonesia.com/nasional/20200422071248-20-495899/tangkal-corona-psbb-bandung-raya-mulai-berlaku-hari-ini)
- [Surabaya PSBB 28 April](https://surabaya.liputan6.com/read/4236093/pelaksanaan-psbb-surabaya-raya-mulai-berlaku-28-april-2020)
- [Jawa Barat PSBB 6 Mei](https://tirto.id/psbb-jawa-barat-mulai-6-mei-ratusan-titik-dijaga-buat-cegah-mudik-fk1M)

#### Data source

Indonesia:

- @kawalCOVID19
- https://covid19.go.id/

Jakarta:

- https://corona.jakarta.go.id/id/data-pemantauan

Internasional:

[Worldmeters](https://www.worldometers.info/coronavirus/#countries)

#### Source Code

[github](https://github.com/dioariadi/covid-19)

#### Chart Style/Type

- https://ig.ft.com/coronavirus-chart/?areas=usa&areas=idn&cumulative=1&logScale=1&perMillion=0&values=cases
- https://github.com/VictimOfMaths/COVID-19/blob/master/LAHeatmaps.R
- https://mobile.twitter.com/jburnmurdoch/status/1256711433829703683


